Extending statistical models for source attribution of zoonotic diseases: A study of campylobacteriosis

### 2018-03-19 paper/finalish/article_1023.tex
adds some discussiony stuff

\section{Discussions}
Models that determine the source of human infection, particularly with zoonotic pathogens that originate in animal populations, are of considerable value to public health policy-makers. However, the complexity of model building may increase because of the complex transmission pathways of pathogens, in particular when the model combines the epidemiological and molecular data. But, is it necessary to use more complex genetic models for the source attribution of zoonoses? There is limited research on whether using a simpler model could work as effective as complex genetic models. Here we developed a relatively simple model to estimate the attribution probability for each source of \textit{Campylobacter} infection. This model differs to the asymmetric Island model, it does not model the pathogenic evolution but uses a simple genetic classification such as a sequence type of each isolate to illustrate the sampling distribution of genotypes. With the aim of filling the research gap, our results showed that the simple model can do at least similar work, compared to the asymmetric Island model.

>>>>

\section{Discussions}
Models that determine the source of human infection, particularly with zoonotic pathogens that originate in animal populations, are of considerable value to public health policy makers. However, the complexity of models may increase due to the complex transmission pathways of pathogens, particularly when a model combines epidemiological and molecular data. An outstanding question is whether such complexity is required, or whether a simpler model may work as effectively. Here we developed a relatively simple model to estimate the attribution probability for each source of \textit{Campylobacter} infection. This model differs to the asymmetric Island model, in that it does not model pathogenic evolution, opting instead to use infer the sampling distribution of genotypes directly from the observed count data.

Our results showed that the simple Dirichlet model and the asymmetric Island model showed largely similar results for the final attribution, with both models demonstrating a clear effect of rurality on attribution: cases in rural areas are more likely to have originated from ruminants, while those in main urban centers are more likely to be of poultry origin. While the overall attribution was consistent, it would be expected that the conditional probabilities for a given genotype might differ markedly between models. For those genotypes observed infrequently (or not at all) among the sources, the Dirichlet model has little information, while the asymmeteric Island model can exploit additional information from cases with similar (but not identical) genetic profiles. In the case of MLST data, with just 7 loci, the majority of human cases and source isolates come from a relatively small number of sequence types which are observed often. Thus the Dirichlet model performs well, as it has sufficient observations to estimate the genotype distribution well where the bulk of the data lie. It is only those genotypes that are rarely observed where it performs poorly, but as they are rarely observed, they don't contribute much to the overall attribution. In other circumstances, such as where we have many more than 7 loci, we would have many more rare genotypes, so that the Dirichlet model might provide little useful information. At the extreme example of whole genome MLST (wgMLST) where each isolate might be unique, it would provide essentially no information at all. In these circumstances, however, the asymmetric Island model would be expected to still perform well, assuming that information could be transferred between similar genotypes.

As with any Bayesian analysis, it is of interest to examine the sensitivity of hte results to the choice of prior distributions. We originally used standard normal priors for regression coefficients on the logit scale. We also considered...
